DSCI 445 - Utilizing Machine Learning for Algae Predictions

### 2020-12-08 slides.Rmd
add results slides content and python code

---
title: "DSCI 445 - Utilizing Machine Learning for Algae Predictions"
author: "Dawson Eliasen, Boston Lee, Miles Austin"
date: "12/7/2020"
output:
  slidy_presentation: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(tmap)
library(tmaptools)
library(USAboundaries)
library(sf)
knitr::opts_chunk$set(echo = FALSE)
```

@@

When looking at the amount of hand-collected (in situ) data, we see that in Colorado only has data for the major municipal water bodies like Horsetooth.
```{r, eval=F}
sampled_map <- st_read("sampled_map.csv", geometry_column='geometry')
states <- us_boundaries()
st_as_sf("sampled_map")
co_outline <- states %>%
  filter(name == 'Colorado') %>%
  st_transform(2163)
tmap_mode("plot")
tm_shape(co_outline) + tm_polygons(alpha=0) +
tm_shape(sampled_map) + tm_fill("sampled", id="site_name",
palette=c("red","blue","green")) + tm_layout(frame=F)
```

@@

##

>>>>

---
title: "DSCI 445 - Utilizing Machine Learning for Algae Predictions"
author: "Dawson Eliasen, Boston Lee, Miles Austin"
date: "12/7/2020"
output:
  beamer_presentation: default
  slidy_presentation: default
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE}
library(knitr)
opts_chunk$set(engine.path='venv/bin/python3')
```

```{python, include=FALSE}
from plots import *
```

@@

When looking at the amount of hand-collected (in situ) data, we see that in Colorado only has data for the major municipal water bodies like Horsetooth.

@@

## Methodology

```{python, echo=FALSE}
plot_histograms()
```

## Random forest feature importance

- We were surprised to see `areasqkm` as the most important feature
- `dwl` as the second most important feature makes sense

```{python, echo=FALSE}
plot_feature_importance()
```

## How does the feature subset affect performance?

- Different models behave similarly
- Adding in `areasqkm` resulted in the biggest performance increase

```{python, echo=FALSE}
plot_subset_performance()
```

## How does the merge threshold affect performance?

- Different models behave similarly
- Interestingly, a merge threshold of 1 is worse than 0 and 2

```{python, echo=FALSE}
plot_thresh_performance()
```

## Results summary

- Our best model was a random forest with a merge threshold of 3 days and included all available features
- Our best model received a CV MSE of about 1.75 and a test set MSE of about 1.09

## Predicted values

```{python, echo=FALSE}
plot_predicted_values()
```
